A cascaded supervised learning approach to inverse reinforcement learning

### 2013-03-29 piot2012cascading/ECML13_cascading.tex
Suites modifs pour JFPDA/ECML

Given a dataset $D_C = \{(s_i,a_i=\pi_E(s_i))_i\}$ of actions $a_i$ (deterministically) chosen by the expert on states $s_i$, we train such a classifier. The classification policy $\pi_C$ is not the end product we are looking for (that would be mere imitation of the expert, not IRL). What is of particular interest to us is the score function $q$ itself. One can easily notice the similarity between Eq.~\eqref{eq:greedy2} and Eq.~\eqref{eq:greedy} that describes the relation between the optimal policy in an MDP and its optimal action-value function. The score function $q$ of the classifier can thus be viewed as some kind of optimal action-value function for the classifier policy $\pi_C$. By inversing the Bellman equation Eq.~\eqref{eq:bellman1} with $q$ in lieu of $Q^\pi_R$, one gets $R^C$, the reward function relative to our score/action-value function $q$:
\begin{equation}
  \label{eq:rc}
  R^C(s,a) =q(s,a) - \sum_{\mathrm{s'}}P(s'|s,a)q(s',\pi_C(s')).
\end{equation}
As we wish to solve the general, or approximate, IRL problem where the transition probabilities $P$ are unknown, our reward function $R^C$ will be approximated with the help of information gathered by interacting with the system. We assume that another dataset $D_R = \{(s_j,a_j,s'_j)_j\}$ is available where $s'_j$ is the state an agent taking action $a_j$ in state $s_j$ transitioned to. Action $a_j$ need not be chosen by any particular policy. The dataset $D_R$ brings us information about the dynamics of the system. From it, we construct datapoints
\begin{equation}
  \label{eq:rj}
  \{\hat r_j = q(s_j,a_j) - \gamma q(s'_j,\pi_C(s'_j))\}_j.
\end{equation}
As $s'_j$ is sampled accordingly to $P(\cdot|s_j,a_j)$ the constructed datapoints help building a good approximation of $R^C(s=s_j,a=a_j)$. A regressor (a simple least-square approximator can do but other solutions could also be envisionned, see Sec.~\ref{subsubsec:subroutines}) is then fed the datapoints $((s_j,a_i),\hat r_j)$ to obtain $\hat R^C$, a generalisation of $\{((s_j,a_j),\hat r_j)_j\}$ over the whole state-action space. The complete algorithm is given in Alg.~\ref{algo:cascading}.

There is no particular constraint on $D_C$ and $D_R$. Evidently, there is a direct link between various qualities of those two sets (amount of data, statistical representativity, etc.) and the classification and regression error. The exact nature of the relationship between these quantities depends on which classifier and regressor are chosen. The theoretical analysis of Sec.\ref{sec:analysis} abstracts itself from the choice of a regressor and a classifier and from the composition of $D_C$ and $D_R$ by reasoning with the classification and regression error. In Sec.\ref{sec:experiments}, %FIXME subsec heuristic
the use of a single dataset to create both $D_C$ and $D_R$ is thoroughly explored.
\begin{algorithm}%[H]
    %\small
  \caption{CSI algorithm}
  \label{algo:cascading}
  \emph{\textbf{Given}} a training set $D_C=\{(s_i,a_i=\pi_E(s_i))\}_{1\leq i \leq D}$ and another training set $D_R=\{(s_{j},a_{j},s'_{j})\}_{1\leq j \leq D'}$\;\\
  \emph{\textbf{Train}} a score function-based classifier on $D_C$, obtaining decision rule $\pi_C$ and score function $q:S\times A \rightarrow \mathbb R$\;\\
  \emph{\textbf{Learn}} a reward function $\hat R^C$ from the dataset $\{((s_{j},a_{j}),\hat{r}_j)\}_{1\leq j \leq D'}$, $\forall (s_j,a_j,s'_j) \in D_R,\hat{r}_j=q(s_{j},a_{j})-\gamma q(s'_{j},\pi_C(s'_{j}))$\;\\
  \emph{\textbf{Output}} the reward function $\hat R^{C}$ \;
\end{algorithm}

Cascading two supervised approaches like we do is a way to inject the MDP structure in the resolution of the problem. Indeed, mere classification only takes into account information from the expert (i.e., which action goes with which state) whereas using the Bellman equation in the expression of $\hat r_j$ makes use of the information lying in the transitions $(s_j,a_j,s'_j)$, namely information about the transition probabilities $P$. The final regression step is a way to generalize this information about $P$ to the whole state-action space in order to have a well-behaved reward function.

@@

In this section, we prove that the deterministic expert policy $\pi_E$ is near optimal for the reward $\hat{R}^C$ the regression step outputs.
More formally, recalling from Sec.~\ref{sec:background} that $\rho_E$ is the stationary distribution of the expert policy, we prove that $\E_{s\sim\rho_E}[V^*_{\hat{R}^C}(s)-V^{\pi_E}_{\hat{R}^C}(s)]$ is bounded by a term that depends on:
\begin{itemize}
  \item the classification error defined as

@@

\epsilon_C=\E_{s\sim\rho_E}[\mathds{1}_{\{\pi_C(s)\neq\pi_E(s)\}}];

@@

\item the regression error defined as 

@@

    \epsilon_R=\max_{\pi\in A^S}\|\epsilon_\pi^R\|_{1,\rho_E},

@@

  with:
  \begin{itemize}
  \item the subscript notation already used for $R_\pi$ and $P_\pi$ in Sec.~\ref{sec:background} meaning that, given an $X\in\mathbb{R}^{S\times A}$, $\pi\in A^S$, and $a\in A$,  $X_\pi\in\mathbb{R}^S$ and $X_a\in\mathbb{R}^S$ are respectively such that: $\forall s\in S, X_\pi(s)=X(s,\pi(s))$ and $\forall s\in S, X_a(s)=X(s,a)$ ;
  \item $\epsilon^R_\pi=R^C_\pi-\hat{R}^C_\pi$ ;
  \item $\|.\|_{1,\mu}$ the $\mu$-weighted $L_1$ norm: $\|f\|_{1,\mu} = \E_{x\sim \mu}[|f(x)|]$, one property of which is that $\forall X, \mu^TX \leq \|X\|_{1,\mu}$;
  \end{itemize}
\item the concentration coefficient $C_* = C_{\hat \pi_C}$ with:
  \begin{itemize}
  \item $C_{\pi}=(1-\gamma)\sum_{t\geq0}\gamma^tc_{\pi}(t), \text{ with } c_{\pi}(t)=\max_{s\in S}\frac{(\rho_E^TP^t_\pi)(s)}{\rho_E(s)}$ ;
  \item $\hat \pi_C$, the optimal policy for the reward $\hat R^C$ output by the algorithm ;
  \end{itemize}
  The constant $C_*$ can be estimated \emph{a posteriori} (after $\hat{R}^C$ is computed). {\it A priori}, $C_*$ can be upper bound by a more usual and general concentration coefficient but $C_*$ gives a tighter final result: one can informally see $C_*$ as a measure of the similarity between the distributions indeuced by $\hat\pi_C$ and $\pi_E$ (roughly, if $\hat{\pi}_C\approx\pi_E$ then $C_* \approx 1$).
\item $\Delta q$, defined by 

@@

    \Delta q &= &\max_{s\in S}(\max_{a\in A}q(s,a)-\min_{a\in A}q(s,a))=\max_{s\in S}(q(s,\pi_C(s))-\min_{a\in A}q(s,a)),
    \end{eqnarray} which could be normalized to $1$ without loss of generality. The range of variation of $q$ induces the one of $R^C$, $\hat{R}^C$ and $V^{\pi_C}_{\hat{R^C}}$.
  \end{itemize}
Readers familiar with the work presented in \cite{klein2012scirl} will see some similarities between the theoretical analyses of SCIRL and CSI. As both algorithms solve the same problem, both theorems naturally bound the same criterion. Another shared feature is the use of the score function $q$ of the classifier as a proxy for the action-value function of the expert $Q^{\pi_E}_{R_E}$. The same structure can therefore be found in both analyses as there is only so many ways to study error propagation in IRL algorithms.\\
The attentive reader, however, will percieve that similarities stop there. The error terms occurring in the two bounds are not related to one another. As CSI makes no use of the feature expectation of the expert, what is known as $\bar\epsilon_Q$ in \cite{klein2012scirl} does not appear in this analysis. Likewise, the regression error $\epsilon_R$ of this paper does not appear in the analysis of SCIRL, which does not use a regressor. Perhaps more subtly, the classification error and classification policy known in both papers as $\epsilon_C$ and $\pi_C$ are not the same. The classification policy of SCIRL is no tantamount to what is called $\pi_C$ here. For SCIRL, $\pi_C$ is the greedy policy for an approximation of the value function of the expert with respect to the reward output by the algorithm. For CSI, $\pi_C$ is the decision rule of the classifier, an object that is not aware of the structure of the MDP.
\begin{theorem}
\label{thm}
Let $\pi_E$ be the deterministic expert policy, $\rho_E$ its stationary distribution and $\hat{R}^C$ the reward the cascading algorithm outputs. We have:

@@

0\leq\E_{s\sim\rho_E}[V^*_{\hat{R}^C}(s)-V^{\pi_E}_{\hat{R}^C}(s)]\leq \frac{1}{1-\gamma}\left(\epsilon_C\Delta q +\epsilon_R(1+C_*)\right).

@@

\end{theorem}

@@

First let's recall some notations, $q\in\mathbb{R}^{S\times A}$ is the score function output by the classification step, $\pi_C$ is a deterministic classifier policy so that $\forall s \in S,\pi_C(s)\in\argmax_{a\in A}q(s,a)$, $R^C\in\mathbb{R}^{S\times A}$ is:
\forall (s,a)\in S\times A, R^C(s,a)=q(s,a) -\gamma\sum_{s'\in S}P(s'|s,a)q(s',\pi_C(s')),
and $\hat{R}^C\in\mathbb{R}^{S\times A}$ is the reward function output by the regression step.

We can now introduce the following vectorial equations:
\begin{align}
&R^C_a=q_a-\gamma P_aq_{\pi_C},
\\
&R^E_a=q_a-\gamma P_aq_{\pi_E},
\\
&\epsilon^R_a=R^C_a-\hat{R}^C_a.
\end{align}
Now, we are going to upper bound the term: $\E_{s\sim\rho_E}[V^*_{\hat{R}^C}-V^{\pi_E}_{\hat{R}^C}]\geq0$ (the lower bound is obvious as $V^*$ is optimal).
Recall that $\hat{\pi}_C$ is a deterministic optimal policy of the reward $\hat{R}^C$. First, the term $V^*_{\hat{R}^C}-V^{\pi_E}_{\hat{R}^C}$ is decomposed:
\begin{equation}
V^*_{\hat{R}^C}-V^{\pi_E}_{\hat{R}^C}=(V^{\hat{\pi}_C}_{\hat{R}^C}-V^{\hat{\pi}_C}_{R^C})+(V^{\hat{\pi}_C}_{R^C}-V^{\pi_E}_{R^C})+(V^{\pi_E}_{R^C}-V^{\pi_E}_{\hat{R}^C}).
\end{equation}
We are going to bound each of these three terms. First, let $\pi$ be a given deterministic policy. We have, using $\epsilon^R = R^C - \hat R^C$:
\begin{equation}
V^{\pi}_{R^C}-V^{\pi}_{\hat{R}^C}=V^{\pi}_{\epsilon^R}=(I-\gamma P_\pi)^{-1}\epsilon^R_{\pi}.
\end{equation}
If $\pi=\pi_E$, we have, thanks to the power series expression of $(I-\gamma P_{\pi_E})^{-1}$ and the definition of $\rho_E$:
\begin{equation}
  \rho_E^T(V^{\pi}_{R^C}-V^{\pi}_{\hat{R}^C})=\rho_E^T(I-\gamma P_{\pi_E})^{-1}\epsilon^R_{\pi_E}=\frac{1}{1-\gamma}\rho_E^T\epsilon^R_{\pi_E}\leq\frac{1}{1-\gamma}\|\epsilon^R_{\pi_E}\|_{1,\rho_E}.\end{equation}

If $\pi\neq\pi_E$, we use the concentration coefficient $C_\pi$. We have then:
\begin{equation}
  \rho_E^T(V^{\pi}_{R^C}-V^{\pi}_{\hat{R}^C})\leq\frac{C_\pi}{1-\gamma}\rho_E^T\epsilon^R_{\pi}\leq\frac{C_\pi}{1-\gamma}\|\epsilon^R_{\pi}\|_{1,\rho_E}.
  \end{equation}
So, using the notations introduced before we stated the theorem, we are able to give an upper bound to the first and third terms (recall also the notation $C_{\hat \pi_C} = C_*$):
\begin{equation}
\rho_E^T((V^{\pi_E}_{R^C}-V^{\pi_E}_{\hat{R}^C})+(V^{\hat{\pi}_C}_{\hat{R}^C}-V^{\hat{\pi}_C}_{R^C}))\leq\frac{1+C_*}{1-\gamma}\epsilon_R.
\end{equation}
Now, there is still an upper bound to find for the second term. It is possible to decompose it as follows:
\begin{equation}
V^{\hat{\pi}_C}_{R^C}-V^{\pi_E}_{R^C}=(V^{\hat{\pi}_C}_{R^C}-V^{\pi_C}_{R^C})+(V^{\pi_C}_{R^C}-V^{\pi_E}_{R^E})+(V^{\pi_E}_{R^E}-V^{\pi_E}_{R^C}).
\end{equation}
By construction, $\pi_C$ is optimal for $R^C$, so $V^{\hat{\pi}_C}_{R^C}-V^{\pi_C}_{R^C}\leq0$ which implies:
\begin{equation}
V^{\hat{\pi}_C}_{R^C}-V^{\pi_E}_{R^C}\leq(V^{\pi_C}_{R^C}-V^{\pi_E}_{R^E})+(V^{\pi_E}_{R^E}-V^{\pi_E}_{R^C}).
\end{equation}
By construction, we have $V^{\pi_C}_{R^C}=q_{\pi_C}$ and $V^{\pi_E}_{R^E}=q_{\pi_E}$, thus:
\begin{eqnarray}
\rho_E^T(V^{\pi_C}_{R^C}-V^{\pi_E}_{R^E})&=&\rho_E^T(q_{\pi_C}-q_{\pi_E})\\
&=&\sum_{s\in S}\rho_E(s)(q(s,\pi_C(s))-q(s,\pi_E(s)))[\mathds{1}_{\{\pi_C(s)\neq\pi_E(s)\}}].
\end{eqnarray}
Using $\Delta q$, we have:
\begin{equation}
\rho_E^T(V^{\pi_C}_{R^C}-V^{\pi_E}_{R^E})\leq\Delta q\sum_{s\in S}\rho_E(s)[\mathds{1}_{\{\pi_C(s)\neq\pi_E(s)\}}]= \Delta q \epsilon_C.
\end{equation}
Finally, we also have:
\begin{align}
\rho_E^T(V^{\pi_E}_{R^E}-V^{\pi_E}_{R^C})&=\rho_E^T(I-\gamma P_{\pi_E})^{-1}(R^E_{\pi_E}-R^C_{\pi_E}),
\\
&=\rho_E^T(I-\gamma P_{\pi_E})^{-1}\gamma P_{\pi_E}(q_{\pi_C}-q_{\pi_E}),
\\
&=\frac{\gamma}{1-\gamma}\rho_E^T(q_{\pi_C}-q_{\pi_E})\leq \frac{\gamma}{1-\gamma}\Delta q \epsilon_C \leq  \frac{1}{1-\gamma}\Delta q \epsilon_C\textrm{ as }\gamma<1.
\end{align}
So the upper bound for the second term is: $\rho_E^T(V^{\hat{\pi}_C}_{R^C}-V^{\pi_E}_{R^C})\leq \frac{\Delta q}{1-\gamma}\epsilon_C$.
If we combine all of the results, we obtain the final bound:
\begin{equation}
0\leq\E_{s\sim\rho_E}[V^*_{\hat{R}^C}(s)-V^{\pi_E}_{\hat{R}^C}(s)]\leq \frac{1}{1-\gamma}(\epsilon_C\Delta q +\epsilon_R(1+C_*))).
\end{equation}
\qed
\end{proof}
A direct corollary of this theorem is that, given a perfect classifier and regressor, CSI produces a reward function for which $\pi_E$ is the \emph{unique} optimal policy.
\begin{corollary}
  \label{cor}
  Assume that $\rho_E>0$ and that the classifier and the regressor are perfect (that is $\epsilon_C = 0$ and $\epsilon_R = 0$). Then $\pi_E$ is the \emph{unique} optimal policy for $\hat R^C$.
\end{corollary}
\begin{proof}

@@

The only constraints the bound of theorem~\ref{thm} implies on datasets $D_R$ and $D_C$ is that they provide enough information to the supervised algorithms to keep both error terms $\epsilon_C$ and $\epsilon_R$ low. In Sec.~\ref{sec:experiments} we deal with a lack of data in dataset $D_R$. We address the problem with the use of heuristics (Sec.~\ref{subsubsec:heuristics}) in order to show the behavior of the CSI algorithm in somewhat more realistic (but difficult) conditions.

More generally, the error terms  $\epsilon_C$ and $\epsilon_R$ can be reduced by a wise choice for the classification and regression subroutines. The litterature is wide enough for methods accomodating most of use cases (lack of data, fast computation, bias/variance trade-off, etc.) to be found. Being able to leverage such common algorithms as multi-class classifiers and regressors is a big advantage of our cascading approach over existing IRL algorithms.

@@

AL via classification has been proposed in for example \cite{ratliff2007imitation} with the help of a structured margin method. Using the non trivial notion of metric in an MDP, the authors of \cite{melo2010learning} build a kernel which is used
in a classification algorithm, showing improvements compared to a
non-structured kernel.

>>>>

Given a dataset $D_C = \{(s_i,a_i=\pi_E(s_i))_i\}$ of actions $a_i$ (deterministically) chosen by the expert on states $s_i$, we train such a classifier. The classification policy $\pi_C$ is not the end product we are looking for (that would be mere supervised imitation of the expert, not IRL). What is of particular interest to us is the score function $q$ itself. One can easily notice the similarity between Eq.~\eqref{eq:greedy2} and Eq.~\eqref{eq:greedy} that describes the relation between the optimal policy in an MDP and its optimal action-value function. The score function $q$ of the classifier can thus be viewed as some kind of optimal action-value function for the classifier policy $\pi_C$. By inversing the Bellman equation \eqref{eq:bellman1} with $q$ in lieu of $Q^\pi_R$, one gets $R^C$, the reward function relative to our score/action-value function $q$:
\begin{equation}
  \label{eq:rc}
  R^C(s,a) =q(s,a) - \sum_{\mathrm{s'}}P(s'|s,a)q(s',\pi_C(s')).
\end{equation}
As we wish to approximately solve the general IRL problem where the transition probabilities $P$ are unknown, our reward function $R^C$ will be approximated with the help of information gathered by interacting with the system. We assume that another dataset $D_R = \{(s_j,a_j,s'_j)_j\}$ is available where $s'_j$ is the state an agent taking action $a_j$ in state $s_j$ transitioned to. Action $a_j$ need not be chosen by any particular policy. The dataset $D_R$ brings us information about the dynamics of the system. From it, we construct datapoints
\begin{equation}
  \label{eq:rj}
  \{\hat r_j = q(s_j,a_j) - \gamma q(s'_j,\pi_C(s'_j))\}_j.
\end{equation}
As $s'_j$ is sampled accordingly to $P(\cdot|s_j,a_j)$ the constructed datapoints help building a good approximation of $R^C(s_j,a_j)$. A regressor (a simple least-square approximator can do but other solutions could also be envisionned, see Sec.~\ref{subsubsec:subroutines}) is then fed the datapoints $((s_j,a_i),\hat r_j)$ to obtain $\hat R^C$, a generalisation of $\{((s_j,a_j),\hat r_j)_j\}$ over the whole state-action space. The complete algorithm is given in Alg.~\ref{algo:cascading}.

There is no particular constraint on $D_C$ and $D_R$. Clearly, there is a direct link between various qualities of those two sets (amount of data, statistical representativity, etc.) and the classification and regression errors. The exact nature of the relationship between these quantities depends on which classifier and regressor are chosen. The theoretical analysis of Sec.~\ref{sec:analysis} abstracts itself from the choice of a regressor and a classifier and from the composition of $D_C$ and $D_R$ by reasoning with the classification and regression errors. In Sec.~\ref{sec:experiments}, %FIXME subsec heuristic
the use of a single dataset to create both $D_C$ and $D_R$ is thoroughly explored.
\begin{algorithm}%[H]
    %\small
  \caption{CSI algorithm}
  \label{algo:cascading}
  \emph{\textbf{Given}} a training set $D_C=\{(s_i,a_i=\pi_E(s_i))\}_{1\leq i \leq D}$ and another training set $D_R=\{(s_{j},a_{j},s'_{j})\}_{1\leq j \leq D'}$\;\\
  \emph{\textbf{Train}} a score function-based classifier on $D_C$, obtaining decision rule $\pi_C$ and score function $q:S\times A \rightarrow \mathbb R$\;\\
  \emph{\textbf{Learn}} a reward function $\hat R^C$ from the dataset $\{((s_{j},a_{j}),\hat{r}_j)\}_{1\leq j \leq D'}$, $\forall (s_j,a_j,s'_j) \in D_R,\hat{r}_j=q(s_{j},a_{j})-\gamma q(s'_{j},\pi_C(s'_{j}))$\;\\
  \emph{\textbf{Output}} the reward function $\hat R^{C}$ \;
\end{algorithm}

Cascading two supervised approaches like we do is a way to inject the MDP structure into the resolution of the problem. Indeed, mere classification only takes into account information from the expert (i.e., which action goes with which state) whereas using the Bellman equation in the expression of $\hat r_j$ makes use of the information lying in the transitions $(s_j,a_j,s'_j)$, namely information about the transition probabilities $P$. The final regression step is a way to generalize this information about $P$ to the whole state-action space in order to have a well-behaved reward function.

@@

In this section, we prove that the deterministic expert policy $\pi_E$ is near optimal for the reward $\hat{R}^C$ the regression step outputs.
More formally, recalling from Sec.~\ref{sec:background} that $\rho_E$ is the stationary distribution of the expert policy, we prove that $\E_{s\sim\rho_E}[V^*_{\hat{R}^C}(s)-V^{\pi_E}_{\hat{R}^C}(s)]$ is bounded by a term that depends on:
\begin{itemize}
  \item the classification error defined as
\begin{equation}  
\epsilon_C=\E_{s\sim\rho_E}[\mathds{1}_{\{\pi_C(s)\neq\pi_E(s)\}}];
\end{equation}
\item the regression error defined as 
  \begin{equation}
    \epsilon_R=\max_{\pi\in A^S}\|\epsilon_\pi^R\|_{1,\rho_E},
  \end{equation}
  with:
  \begin{itemize}
  \item the subscript notation already used for $R_\pi$ and $P_\pi$ in Sec.~\ref{sec:background} meaning that, given an $X\in\mathbb{R}^{S\times A}$, $\pi\in A^S$, and $a\in A$,  $X_\pi\in\mathbb{R}^S$ and $X_a\in\mathbb{R}^S$ are respectively such that: $\forall s\in S, X_\pi(s)=X(s,\pi(s))$ and $\forall s\in S, X_a(s)=X(s,a)$ ;
  \item $\epsilon^R_\pi=R^C_\pi-\hat{R}^C_\pi$ ;
  \item $\|.\|_{1,\mu}$ the $\mu$-weighted $L_1$ norm: $\|f\|_{1,\mu} = \E_{x\sim \mu}[|f(x)|]$;
  \end{itemize}
\item the concentration coefficient $C_* = C_{\hat \pi_C}$ with:
  \begin{itemize}
  \item $C_{\pi}=(1-\gamma)\sum_{t\geq0}\gamma^tc_{\pi}(t), \text{ with } c_{\pi}(t)=\max_{s\in S}\frac{(\rho_E^TP^t_\pi)(s)}{\rho_E(s)}$ ;
  \item $\hat \pi_C$, the optimal policy for the reward $\hat R^C$ output by the algorithm ;
  \end{itemize}
  The constant $C_*$ can be estimated \emph{a posteriori} (after $\hat{R}^C$ is computed). {\it A priori}, $C_*$ can be upper-bounded by a more usual and general concentration coefficient but $C_*$ gives a tighter final result: one can informally see $C_*$ as a measure of the similarity between the distributions induced by $\hat\pi_C$ and $\pi_E$ (roughly, if $\hat{\pi}_C\approx\pi_E$ then $C_* \approx 1$).
\item $\Delta q$, defined by 
  \begin{eqnarray}
    \Delta q &= &\max_{s\in S}(\max_{a\in A}q(s,a)-\min_{a\in A}q(s,a))=\max_{s\in S}(q(s,\pi_C(s))-\min_{a\in A}q(s,a)),
    \end{eqnarray} which could be normalized to $1$ without loss of generality. The range of variation of $q$ induces the one of $R^C$, $\hat{R}^C$ and $V^{\pi_C}_{\hat{R^C}}$.
  \end{itemize}
\begin{theorem}
\label{thm}
Let $\pi_E$ be the deterministic expert policy, $\rho_E$ its stationary distribution and $\hat{R}^C$ the reward the cascading algorithm outputs. We have:
\begin{equation}
0\leq\E_{s\sim\rho_E}[V^*_{\hat{R}^C}(s)-V^{\pi_E}_{\hat{R}^C}(s)]\leq \frac{1}{1-\gamma}\left(\epsilon_C\Delta q +\epsilon_R(1+C_*)\right).
\end{equation}
\end{theorem}
\begin{proof}
First let's recall some notations, $q\in\mathbb{R}^{S\times A}$ is the score function output by the classification step, $\pi_C$ is a deterministic classifier policy so that $\forall s \in S,\pi_C(s)\in\argmax_{a\in A}q(s,a)$, $R^C\in\mathbb{R}^{S\times A}$ is:
\begin{equation}
\forall (s,a)\in S\times A, R^C(s,a)=q(s,a) -\gamma\sum_{s'\in S}P(s'|s,a)q(s',\pi_C(s')),
\end{equation}
and $\hat{R}^C\in\mathbb{R}^{S\times A}$ is the reward function output by the regression step.

We can now introduce the following vectorial equations:
\begin{align}
&R^C_a=q_a-\gamma P_aq_{\pi_C},
\\
&R^E_a=q_a-\gamma P_aq_{\pi_E},
\\
&\epsilon^R_a=R^C_a-\hat{R}^C_a.
\end{align}
Now, we are going to upper bound the term: $\E_{s\sim\rho_E}[V^*_{\hat{R}^C}-V^{\pi_E}_{\hat{R}^C}]\geq0$ (the lower bound is obvious as $V^*$ is optimal).
Recall that $\hat{\pi}_C$ is a deterministic optimal policy of the reward $\hat{R}^C$. First, the term $V^*_{\hat{R}^C}-V^{\pi_E}_{\hat{R}^C}$ is decomposed:
\begin{equation}
V^*_{\hat{R}^C}-V^{\pi_E}_{\hat{R}^C}=(V^{\hat{\pi}_C}_{\hat{R}^C}-V^{\hat{\pi}_C}_{R^C})+(V^{\hat{\pi}_C}_{R^C}-V^{\pi_E}_{R^C})+(V^{\pi_E}_{R^C}-V^{\pi_E}_{\hat{R}^C}).
\end{equation}
We are going to bound each of these three terms. First, let $\pi$ be a given deterministic policy. We have, using $\epsilon^R = R^C - \hat R^C$:
\begin{equation}
V^{\pi}_{R^C}-V^{\pi}_{\hat{R}^C}=V^{\pi}_{\epsilon^R}=(I-\gamma P_\pi)^{-1}\epsilon^R_{\pi}.
\end{equation}
If $\pi=\pi_E$, we have, thanks to the power series expression of $(I-\gamma P_{\pi_E})^{-1}$, the definition of $\rho_E$ and the definition of the $\mu$-weighted $L_1$ norm, one property of which is that $\forall X, \mu^TX \leq \|X\|_{1,\mu}$:
\begin{equation}
  \rho_E^T(V^{\pi}_{R^C}-V^{\pi}_{\hat{R}^C})=\rho_E^T(I-\gamma P_{\pi_E})^{-1}\epsilon^R_{\pi_E}=\frac{1}{1-\gamma}\rho_E^T\epsilon^R_{\pi_E}\leq\frac{1}{1-\gamma}\|\epsilon^R_{\pi_E}\|_{1,\rho_E}.\end{equation}

If $\pi\neq\pi_E$, we use the concentration coefficient $C_\pi$. We have then:

@@

  \rho_E^T(V^{\pi}_{R^C}-V^{\pi}_{\hat{R}^C})\leq\frac{C_\pi}{1-\gamma}\rho_E^T\epsilon^R_{\pi}\leq\frac{C_\pi}{1-\gamma}\|\epsilon^R_{\pi}\|_{1,\rho_E}.

@@

So, using the notations introduced before we stated the theorem, we are able to give an upper bound to the first and third terms (recall also the notation $C_{\hat \pi_C} = C_*$):

@@

\rho_E^T((V^{\pi_E}_{R^C}-V^{\pi_E}_{\hat{R}^C})+(V^{\hat{\pi}_C}_{\hat{R}^C}-V^{\hat{\pi}_C}_{R^C}))\leq\frac{1+C_*}{1-\gamma}\epsilon_R.

@@

Now, there is still an upper bound to find for the second term. It is possible to decompose it as follows:
V^{\hat{\pi}_C}_{R^C}-V^{\pi_E}_{R^C}=(V^{\hat{\pi}_C}_{R^C}-V^{\pi_C}_{R^C})+(V^{\pi_C}_{R^C}-V^{\pi_E}_{R^E})+(V^{\pi_E}_{R^E}-V^{\pi_E}_{R^C}).
By construction, $\pi_C$ is optimal for $R^C$, so $V^{\hat{\pi}_C}_{R^C}-V^{\pi_C}_{R^C}\leq0$ which implies:
V^{\hat{\pi}_C}_{R^C}-V^{\pi_E}_{R^C}\leq(V^{\pi_C}_{R^C}-V^{\pi_E}_{R^E})+(V^{\pi_E}_{R^E}-V^{\pi_E}_{R^C}).
By construction, we have $V^{\pi_C}_{R^C}=q_{\pi_C}$ and $V^{\pi_E}_{R^E}=q_{\pi_E}$, thus:

@@

\rho_E^T(V^{\pi_C}_{R^C}-V^{\pi_E}_{R^E})&=&\rho_E^T(q_{\pi_C}-q_{\pi_E})\\
&=&\sum_{s\in S}\rho_E(s)(q(s,\pi_C(s))-q(s,\pi_E(s)))[\mathds{1}_{\{\pi_C(s)\neq\pi_E(s)\}}].
\end{eqnarray}
Using $\Delta q$, we have:

@@

\rho_E^T(V^{\pi_C}_{R^C}-V^{\pi_E}_{R^E})\leq\Delta q\sum_{s\in S}\rho_E(s)[\mathds{1}_{\{\pi_C(s)\neq\pi_E(s)\}}]= \Delta q \epsilon_C.

@@

Finally, we also have:
\begin{align}
\rho_E^T(V^{\pi_E}_{R^E}-V^{\pi_E}_{R^C})&=\rho_E^T(I-\gamma P_{\pi_E})^{-1}(R^E_{\pi_E}-R^C_{\pi_E}),
\\
&=\rho_E^T(I-\gamma P_{\pi_E})^{-1}\gamma P_{\pi_E}(q_{\pi_C}-q_{\pi_E}),
\\
&=\frac{\gamma}{1-\gamma}\rho_E^T(q_{\pi_C}-q_{\pi_E})\leq \frac{\gamma}{1-\gamma}\Delta q \epsilon_C \leq  \frac{1}{1-\gamma}\Delta q \epsilon_C\textrm{ as }\gamma<1.
\end{align}
So the upper bound for the second term is: $\rho_E^T(V^{\hat{\pi}_C}_{R^C}-V^{\pi_E}_{R^C})\leq \frac{\Delta q}{1-\gamma}\epsilon_C$.
If we combine all of the results, we obtain the final bound:
0\leq\E_{s\sim\rho_E}[V^*_{\hat{R}^C}(s)-V^{\pi_E}_{\hat{R}^C}(s)]\leq \frac{1}{1-\gamma}(\epsilon_C\Delta q +\epsilon_R(1+C_*))).
\qed
\end{proof}
Readers familiar with the work presented in \cite{klein2012scirl} will see some similarities between the theoretical analyses of SCIRL and CSI as both study error propagation in IRL algorithms. Another shared feature is the use of the score function $q$ of the classifier as a proxy for the action-value function of the expert $Q^{\pi_E}_{R_E}$. \\
The attentive reader, however, will percieve that similarities stop there. The error terms occurring in the two bounds are not related to one another. As CSI makes no use of the feature expectation of the expert, what is known as $\bar\epsilon_Q$ in \cite{klein2012scirl} does not appear in this analysis. Likewise, the regression error $\epsilon_R$ of this paper does not appear in the analysis of SCIRL, which does not use a regressor. Perhaps more subtly, the classification error and classification policy known in both papers as $\epsilon_C$ and $\pi_C$ are not the same. The classification policy of SCIRL is no tantamount to what is called $\pi_C$ here. For SCIRL, $\pi_C$ is the greedy policy for an approximation of the value function of the expert with respect to the reward output by the algorithm. For CSI, $\pi_C$ is the decision rule of the classifier, an object that is not aware of the structure of the MDP. We shall also mention that the error terms appearing in the CSI bound are more standard than the ones of SCIRL ({\it e.g.}, regression error vs feature expectation estimation error) thus they may be easier to control.
A direct corollary of this theorem is that, given perfect classifier and regressor, CSI produces a reward function for which $\pi_E$ is the \emph{unique} optimal policy.
\begin{corollary}
  \label{cor}
  Assume that $\rho_E>0$ and that the classifier and the regressor are perfect ($\epsilon_C = 0$ and $\epsilon_R = 0$). Then $\pi_E$ is the \emph{unique} optimal policy for $\hat R^C$.
\end{corollary}

@@

@@

The only constraints the bound of Th.~\ref{thm} implies on datasets $D_R$ and $D_C$ is that they provide enough information to the supervised algorithms to keep both error terms $\epsilon_C$ and $\epsilon_R$ low. In Sec.~\ref{sec:experiments} we deal with a lack of data in dataset $D_R$. We address the problem with the use of heuristics (Sec.~\ref{subsubsec:heuristics}) in order to show the behavior of the CSI algorithm in somewhat more realistic (but difficult) conditions.

More generally, the error terms  $\epsilon_C$ and $\epsilon_R$ can be reduced by a wise choice for the classification and regression algorithms. The litterature is wide enough for methods accomodating most of use cases (lack of data, fast computation, bias/variance trade-off, etc.) to be found. Being able to leverage such common algorithms as multi-class classifiers and regressors is a big advantage of our cascading approach over existing IRL algorithms.

@@

AL via classification has been proposed for example in \cite{ratliff2007imitation}, with the help of a structured margin method. Using the non trivial notion of metric in an MDP, the authors of \cite{melo2010learning} build a kernel which is used
in a classification algorithm, showing improvements compared to a
non-structured kernel.
